---
title: "pPaper"
author: "Dylan Sun"
date: "2/16/2017"
output: pdf_document
---

### Load packages
```{r, message=FALSE, warning=FALSE}
library(data.table)
library(readxl)
library(survival)
library(knitr)
library(tableone)

```

### Load data
```{r}
cardiac <- read_excel("~/Downloads/cardiac_mirna.xlsx")
cardiac <- data.table(cardiac, na.strings = c("NA"))
library(readr)
write_tsv(cardiac, "cardiac.tsv")
cardiac <- fread("cardiac.tsv")

cardiac <- cardiac[, PTV_Volume..cc. := as.numeric(PTV_Volume..cc.)]
cardiac <- cardiac[, Framingham.Coronary.Heart.Disease.Risk.Score..MDCalc. :=
                     as.numeric(Framingham.Coronary.Heart.Disease.Risk.Score..MDCalc.)]
setnames(cardiac, c("Grade 3+ Cardiac Event", "Grade 2+ Cardiac Event", "Event.of.Death..0.no.1.yes."), c("Grade3", "Grade2", "EventOfDeath"))

### add column: whether or not the person died BEFORE having a grade 2 event
cardiac <- cardiac[, ]
```

Number of people that died from cardiac disease?:
```{r}
dim(cardiac[`Grade 2+ Cardiac Event` == 0 & Event.of.Death..0.no.1.yes. == 1])
dim(cardiac[`Grade 3+ Cardiac Event` == 0 & Event.of.Death..0.no.1.yes. == 1])
```


```{r}
# percentages are calculated after excluding missing values
vars <- c("Age", "KPS", "Total.Chol..btw.130.320..", "HDL.Chol.Found..0.no.1.yes.",
          "HDL.Chol..btw.20.100..", "Systolic.BP..btw.20.190..", "Diastolic.BP..btw.30.140..",
          "Framingham.Coronary.Heart.Disease.Risk.Score..MDCalc.", "Final.Dose(tumor)", 
          "BED.Final.Dose(tumor)", "survival", "Volume Heart (cc)", 
          "heart_Meandose", "heart_v5", "heart_V10", "heart_V20", "heart_V30", "heart_V40", 
          "heart_V50", "heart_V60", "D0.5cc.LQ......2.5..EQD2Gy.", "D2cc.LQ......2.5..EQD2Gy.", 
          "Charlson.Comorbidity.Score", "Maximum.Heart.Event.Grade", "time.to.grade3",
          "time.to.grade2", "PTV_Volume..cc.", "Institution", "Gender.M.F", "Race..AA...0.White...1.Other...2", 
              "Current.Smoker.0.no..1.yes", "Smoking.status.0.never...........1..former.2.current",
              "Pack.Year.History", "Diabetes.0.no..1.yes", "High.Cholestorol.Treated..0.no.1.yes", 
              "BP.found..0.no.1.yes.", "Hypertension.0.no.1.yes", 
              "HBP.Treated.0.no..1.yes..on.meds..", "Aspirin.Use.0.no..1.yes", 
              "Pre.Existing.CAD..stent..CABG..acute.coronary.syndrome..0.no.1.yes", 
               "Pre.Existing.Congestive.heart.failure..CHF..0.no.1.yes", 
              "Pre.existing.arrhythmia.0.no..1.yes", 
              "Pre.Existing.Cardiac.Disease..CADorCHF..0.no.1.yes", "Simple.Stage", 
              "Overall.Stage", "T.Stage", "N.Stage", "M.Stage", 
              "Concurrent.chemotherapy..0.no.1.yes.", "Event.of.Death..0.no.1.yes.", 
              "Grade 3+ Cardiac Event", "Grade 2+ Cardiac Event",
              "Pre.Existing.Cardiac.Disease..CADorCHF..1.no.0.yes" )
cat_vars <- c("Institution", "Gender.M.F", "Race..AA...0.White...1.Other...2", 
              "Current.Smoker.0.no..1.yes", "Smoking.status.0.never...........1..former.2.current",
              "Diabetes.0.no..1.yes", "High.Cholestorol.Treated..0.no.1.yes", 
              "BP.found..0.no.1.yes.", "Hypertension.0.no.1.yes", 
              "HBP.Treated.0.no..1.yes..on.meds..", "Aspirin.Use.0.no..1.yes", 
              "Pre.Existing.CAD..stent..CABG..acute.coronary.syndrome..0.no.1.yes", 
               "Pre.Existing.Congestive.heart.failure..CHF..0.no.1.yes", 
              "Pre.existing.arrhythmia.0.no..1.yes", 
              "Pre.Existing.Cardiac.Disease..CADorCHF..0.no.1.yes", "Simple.Stage", 
              "Overall.Stage", "T.Stage", "N.Stage", "M.Stage", 
              "Concurrent.chemotherapy..0.no.1.yes.", "Event.of.Death..0.no.1.yes.", 
              "Grade 3+ Cardiac Event", "Grade 2+ Cardiac Event",
              "Pre.Existing.Cardiac.Disease..CADorCHF..1.no.0.yes" )
tab1 <- CreateTableOne(data = cardiac, vars = vars, factorVars = cat_vars, includeNA = F)
tab1mat <- print(tab1, showAllLevels = T)
write.csv(tab1mat, file = "tableone.csv")
descriptive <- fread("tableone.csv")
```


## Important descriptive stats
Be careful; there are two "Pre.Existing.Cardiac.Disease..CADorCHF..0.no.1.yes" variables...
```{r}
# percentages are calculated after excluding missing values
vars <- c("Age", "Framingham.Coronary.Heart.Disease.Risk.Score..MDCalc.",
            "Grade3", "Grade2",
              "Pre.Existing.Cardiac.Disease..CADorCHF..0.no.1.yes" )
cat_vars <- c("Grade3", "Grade2",
              "Pre.Existing.Cardiac.Disease..CADorCHF..0.no.1.yes" )
tab1 <- CreateTableOne(data = cardiac, vars = vars, factorVars = cat_vars, strata = "EventOfDeath")
tab1mat <- print(tab1, showAllLevels = T)
```



## Model 
Variables: Radiation dose, heart disease y/n, framingham risk score (only have his if no cardiac disease)
Outcome: Survival (months) or last follow-up 

Age, gender, other variables are sort of contained in the Framingham risk score (but not for people that have pre-existing disease). There are few non-whites in the study and relatively few females in the study, so including these is questionable. 

Framingham score might be associated with age and gender? Positive correlation pattern between Age and Framingham risk score. Framingham risk score is composed of other variables including:  age, sex, LDL cholesterol, HDL cholesterol, blood pressure, and smoking. Therefore including risk score while also including these other variables is a little suspicious. We only have Framingham risk score if the patient does not have pre-existing cardiac disease. Keep in mind also that the scoring is a bit different between men and women. 

We may have to fit separate models depending on whether or not the patient has a pre-existing cardiac condition (i.e. whether or not the patient has a Framingham risk score; right now when you fit the model it just omits the entries with no risk score, because there's missing data)

```{r}
plot(cardiac$Framingham.Coronary.Heart.Disease.Risk.Score..MDCalc., cardiac$Age)
sum(cardiac$Pre.Existing.Cardiac.Disease..CADorCHF..1.no.0.yes)/dim(cardiac)[1]
```


```{r}
surv_death <- Surv(cardiac[, survival], cardiac[, EventOfDeath])
cox_model <- coxph(surv_death ~  Framingham.Coronary.Heart.Disease.Risk.Score..MDCalc. +
                     heart_Meandose,
                   data = cardiac)
summary(cox_model)
```


```{r}
surv_grade3 <- Surv(time = cardiac[, time.to.grade3], event = cardiac[, Grade3])
cox_model <- coxph(surv_grade3 ~ 
                     Framingham.Coronary.Heart.Disease.Risk.Score..MDCalc. + 
                     heart_Meandose, data = cardiac)
summary(cox_model)
```

```{r}
surv_grade2 <- Surv(time = cardiac[, time.to.grade2], event = cardiac[, Grade2])
cox_model <- coxph(surv_grade2 ~ 
                     Framingham.Coronary.Heart.Disease.Risk.Score..MDCalc. + 
                     heart_Meandose, data = cardiac)
summary(cox_model)
plot(survfit(cox_model))
```


### Look at columns to figure out the competing risk situation
```{r}
View(cardiac[, .(survival, time.to.grade3, time.to.grade2, EventOfDeath, Grade3, Grade2)])
```







